Computing volume via non-linear rectangular prisms

### 2019-08-08 Cav2D/article-template.tex
added initial formula for cav 2d

\section{Two Dimensional Cavalieri Integral}
Let $\mathcal{S}$ be the solid which is bounded by the planes $f(x,y)=-x-y+8$, $2x$, $2(x-4)$, $y$, $y-4$ and the $xy$-plane. Solid $\mathcal{S}$ is depicted 
in Figure~\ref{fig:prismatoid}, Figure~\ref{fig:prismatoid_real} and Figure~\ref{fig:prismatoid_solid}. Solid $\mathcal{S}$ is a prismatoid, since all of its 
vertices lie in two parallel planes. Furthermore, all of its lateral faces are trapezoids or triangles. The volume of $\mathcal{S}$, $V_{\mathcal{S}}$, can easily be computed using the 
volume equation of a prismatoid. The details of this computation is described in the caption of Figure~\ref{fig:prismatoid}. According to the prismatoid volume 
equation $V_{\mathcal{S}}=25.6$.

Note that $\mathbf{c}(t) = -\left<\frac{1}{2}t,t,t\right>$ is the line of intersection between the planes $2x$ and $y$. Furthermore, the line 
$\mathbf{c}(t)$ passes through the origin. The line $\mathbf{c}(t)$ is depicted and labelled in Figure~\ref{fig:prismatoid_solid}. Let 
\begin{equation}
R = \{(x,y)\in\mathbb{R}^2|0 \geq x \geq 4, 0 \geq x \geq 4\} \nonumber
\end{equation}
and
\begin{equation}
S = \{(x,y)\in\mathbb{R}^2|-\sfrac{1}{3}y+\sfrac{8}{3}\geq x\geq -\sfrac{1}{3}y+\sfrac{16}{3}, \sfrac{1}{2} x + 4\geq y \geq \sfrac{1}{2} x + 6 \}. \nonumber
\end{equation}
The region $S$ and $R$ is depicted in Figure~\ref{fig:prismatoid_solid} and Figure~\ref{fig:prismatoid_regions}. Furthermore, let the transformation $T$ map the $xy$--coordinate $(x_i^1,y_i^1)\in R$ at which $\mathbf{c}(t) + \left <x_i^1,y_i^2,0 \right >$ intersects the 
$xy$-plane to the $xy$--coordinate $(x_i^2,y_i^2)\in S$ at which $\mathbf{c}(t) + \left <x_i^1,y_i^2,0 \right >$ intersects $f(x,y)$.

>>>>

\section{Double Cavalieri Integral}
Let $\mathcal{S}$ be the solid which is bounded by the planes $f(x,y)=-x-y+8$, $2x$, $2(x-4)$, $y$, $y-4$ and the $xy$-plane. Solid $\mathcal{S}$ is depicted 
in Figure~\ref{fig:prismatoid}, Figure~\ref{fig:prismatoid_real} and Figure~\ref{fig:prismatoid_solid}. Solid $\mathcal{S}$ is a prismatoid, since all of its 
vertices lie in two parallel planes. Furthermore, all of its lateral faces are trapezoids or triangles. The volume of $\mathcal{S}$, $V_{\mathcal{S}}$, can easily be computed using the 
volume equation of a prismatoid. The details of this computation is described in the caption of Figure~\ref{fig:prismatoid}. According to the prismatoid volume 
equation $V_{\mathcal{S}}=25.6$.

Let $\mathbf{c}(t) = -\left<\frac{1}{2}t,t,t\right>$ be the line of intersection between the planes $2x$ and $y$. Furthermore, note that the line 
$\mathbf{c}(t)$ passes through the origin. The line $\mathbf{c}(t)$ is depicted and labelled in Figure~\ref{fig:prismatoid_solid}. Let 
\begin{equation}
R = \{(x,y)\in\mathbb{R}^2|0 \geq x \geq 4, 0 \geq x \geq 4\} \nonumber
\end{equation}
and
\begin{equation}
S = \{(x,y)\in\mathbb{R}^2|-\sfrac{1}{3}y+\sfrac{8}{3}\leq x\leq -\sfrac{1}{3}y+\sfrac{16}{3}, \sfrac{1}{2} x + 4\leq y \leq \sfrac{1}{2} x + 6 \}. \nonumber
\end{equation}
The region $S$ and $R$ are depicted in Figure~\ref{fig:prismatoid_solid} and Figure~\ref{fig:prismatoid_regions}. Let us now subdivide $R$ into $n\times m$
subrectangles $R_{ij} = [x_{i},x_{i+1}] \times [y_{j},y_{j+1}]$ of equal width $\Delta x^1 = \Delta x_i^1 = x_{i+1}-x_i$ and length $\Delta y^1= \Delta y_j^1 = y_{j+1}-y_j$. 
Furthermore, let the transformation $T$ map the $xy$--coordinate $(x_i^1,y_i^1)\in R$ (the lower left vertice of $R_{ij}$) at which $\mathbf{c}(t) + \left <x_i^1,y_i^1,0 \right >$ intersects the 
$xy$-plane to the $xy$--coordinate $(x_i^2,y_j^2)\in S$ at which $\mathbf{c}(t) + \left <x_i^1,y_j^1,0 \right >$ intersects $f(x,y)$ (Figure~\ref{fig:prismatoid_solid} and Figure~\ref{fig:prismatoid_regions}). 
%The transformation $T$ is defined by the following equations:
%\begin{equation}
%x_i^2 = h(x_i^1,y_i^1)
%\end{equation}
%and
%\begin{equation}
%y_i^2 = s(x_i^1,y_i^1)
%\end{equation}
We can approximate $\mathcal{S}_V$ via:
\begin{equation}
V_{\mathcal{S}} \approx \sum_{i=0}^{n-1} \sum_{j=0}^{m-1} f(x_i^2,y_j^2) \Delta x_i^1 \Delta y_j^1.
\end{equation}
Let us verify that this is indeed the case:
\begin{itemize}
\item $\Delta x_i^1$ and $\Delta y_i^i$ represent the width and the length of subrectangle $R_{ij}$.
\item The product $\Delta A = \Delta x_i^1\times \Delta y_i^i$ represent the area of subrectangle $R_{ij}$.
\item $f(x_i^2,y_i^2)$ represents the height of a non-linear rectangular column or prism $\mathcal{K}_{ij}$ whose base is $R_{ij}$. Furthermore, the lateral edges of this column are all translations 
of $\mathbf{c}(t)$ along the $xy$-plane. 
\item The product $f(x_i^2,y_i^2)\times \Delta A$ represents the volume of $\mathcal{K}_{ij}$ (due to Cavalieri's principle).
\item The double Cavalieri sum $\sum_{i=0}^{n-1} \sum_{j=0}^{m-1} f(x_i^2,y_j^2) \Delta A$, therefore, approximates $V_{\mathcal{S}}$. 
\item In the limit the double Cavalieri sum approaches $V_{\mathcal{S}}$ (if this limit exists). 
\end{itemize}

In this paper, we will refer to aformentioned limit as the two dimensional Cavalieri integral and we will denote this integral as follows:

\begin{equation}
\label{eq:cav_2d}
\iint \limits_{\mathcal{C}} f(x,y)~dA = \lim_{n,m \rightarrow \infty}  \sum_{i=0}^{n-1} \sum_{j=0}^{m-1} f(x_i^2,y_j^2) \Delta x_i^1 \Delta y_j^1,
\end{equation}
with $\mathcal{C} = \{\mathbf{c}(t),R\}$. As is the case for the single Cavalieri integral, Equation~\eqref{eq:cav_2d} is difficult to evaluate in its current form. Luckily, the double 
Cavalieri integral can be converted into an equivalent Riemann integral using the transformation $T$. In general the transformation $T$ is defined via the following 
equations:
\begin{equation}
x_i^2 = h(x_i^1,y_i^2)~\textrm{and}~y_i^2 = s(x_i^1,y_i^1) 
\end{equation}
We can now rewrite Equation~\eqref{eq:cav_2d} as follows:

\begin{eqnarray}
\iint\limits_{\!\mathcal{C}} f(x,y) dA &=& \lim_{n,m\rightarrow \infty} \sum_{j=1}^n\sum_{i=1}^n f(x_{ij}^2,y_{ij}^2) \Delta x_{ij}^1\Delta y_{ij}^1\\
&=&  \lim_{n,m\rightarrow \infty} \sum_{j=1}^n\sum_{i=1}^n f(g(x_{ij}^1,y_{ij}^1),r(x_{ij}^1,y_{ij}^1)) \Delta x_{ij}^1\Delta y_{ij}^1\\
&=& \int_a^b\int_c^d f(g(x,y),r(x,y))~dx~dy.
\end{eqnarray}

Note that, for every coordinate pair there exists a unique $t\in \mathbb{R}$ such that

\begin{equation}
\label{eq:vector_2}
\left< x_i^2, y_i^2, f(x_i^2,y_i^2) \right > = \left < x_i^1, y_i^1, 0 \right > + \left <c_x(t),c_y(t),c_z(t)\right >  
\end{equation}

The above equation can be re-expressed as

\begin{equation}
c_z(t) = f(x_i^1 + c_x(t),y_i^1 + c_y(t)). 
\end{equation}

We can now solve for $t$ in terms of $x_i^1$ and $y_i^1$. If we substitute this solution into Equation~\ref{eq:vector_2} we obtain

\begin{equation}
x_i^2 = h(x_i^1,y_i^1) = x_i^1 + c_x\circ t(x_i^1,y_i^1) 
\end{equation}
and 
\begin{equation}
y_i^2 = s(x_i^1,y_i^1) = y_i^1 + c_y\circ t(x_i^1,y_i^1) 
\end{equation}
